---
title: "Homework 4 Task 2"
author: "Sarah Lam"
date: "3/3/2022"
output: html_document
output: 
  html_document:
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(tsibble)
library(feasts)
library(broom)
library(fable)
library(lubridate)
```
## Overview 

TASK 2: Willamette Falls fish passage time series summary
The goal of this task is to create a single beautiful, customized RMarkdown document to explore time series data in 3 parts, based on this Willamette Falls fish passage data.
Data were shared by and accessed from Columbia River DART (Data Access in Real Time).

```{r}
#read in data
fish_passing <- read_csv(here("data", "willamette_fish_passage.csv")) %>% 
  clean_names()
```

Tab 1: Original time series
A finalized, static graph of adult passage for coho, jack coho, and steelhead salmon (you decide if these all occupy the same panel, or if they are in separate panels). Add a figure caption. Replace NA values with zero. 
2 - 3 bullet points summarizing major patterns / notable trends in the salmon passage time series data
```{r}
#create time series
fish_ts <- fish_passing %>% 
  clean_names() %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  mutate(date = mdy(date)) %>% 
  select(date, coho, jack_coho, steelhead) %>% 
  as_tsibble(key = NULL,index = date)

fish_ts_pivot <- fish_ts %>% 
  pivot_longer(c(coho, jack_coho, steelhead),
               names_to = "species",
               values_to = "count") %>% 
   mutate(species = case_when(
    species == 'coho' ~ "Coho",
    species == 'jack_coho' ~ "Jack Coho",
    species == 'steelhead' ~ "Steelhead"))
```

```{r}
#plot time series 
ggplot(data = fish_ts_pivot,
       aes(x = date,
           y = count)) +
  geom_line(aes(color = species)) +
  labs(x = 'Date',
       y = 'Number of Fish',
       caption = "**Figure 1.** This figure shows fish ladder use at the Willanette Falls ladder by Coho, Jack Coho, and Steelhead as counts by month between 2001 and 2010.",
       title = "Salmon Species Use of Willamette Falls Fish Ladder")
```

## Analysis Conclusions
  - Steelhead are the most consistent users of the fish ladder
  - Coho had the largest ladder us in 2010
  - Jack Coho species is the most infrequent user of the fish ladder, or the data had a hard time capturing their usage.


## Seasonal Analysis
Tab 2: Seasonplots
A finalized seasonplot for each species (coho, jack coho, steelhead) in its own panel. You can either facet_wrap/facet_grid, or use patchwork or cowplot, to make the figure. Add a caption.
Add 2 - 3 bullet points summarizing the major trends you see in the seasonplots.
```{r}
### create a finalized seasonplot for each species in its own panel (patchwork)
#steelhead_season <- salmon_ts %>% 
 # gg_season(y = steelhead,
  #          pal = viridis(10),
   #         show.legend = FALSE)
# coho_season <- salmon_ts %>% 
 # gg_season(y = coho,
    #        pal = viridis(10),
     #       show.legend = FALSE)
# jack_coho_season <- salmon_ts %>% 
  # gg_season(y = jack_coho,
        #    pal = viridis(10)) 
salmon_ts_long %>% 
  gg_season(y = count,
            pal = magma(10)) +
  labs(y = "Amount of Fish",
       x = "Month",
       color = 'Year',
       title = "Salmon Ladder Use by Season",
       caption = "PUT CAPTION HERE (CITATION)")
```

## Analysis Conclusions
  - Steelhead salmon use the fish ladder more from January - July than August - December.
  - Both Jack Coho and Coho only use the ladder in the fall months, specifically around October.
  - None of the 3 salmon species use the ladder frequently during November and December. 

## Annual Species Count
Tab 3: Annual counts by species
A finalized figure of annual totals for fish passage, for each of the 3 species (coho, jack coho, steelhead). You decide if this is in a single panel, or multiple panels. Add a figure caption. 
Add 2 - 3 bullet points summarizing major trends you see in the annual totals by species from 2000 - 2010.
Here’s the general idea for the layout (top “Overview” section, with the three tabs containing your visualizations below): 
```{r}
salmon_ts_year <- 
  salmon %>% 
  janitor::clean_names() %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .))) %>% 
  mutate(date = lubridate::mdy(date),
         year = year(date)) %>% 
    pivot_longer(c(steelhead, coho, jack_coho),
               names_to = "fish",
               values_to = "count") %>% 
  group_by(year, fish) %>% 
  summarize(yearly_count = sum(count))
ggplot(data = salmon_ts_year, aes(x = year, y = yearly_count)) +
  geom_line(aes(color = fish)) +
  scale_fill_manual(values = c('purple4', 'blue', 'salmon')) +
  scale_x_continuous(breaks = c(2001, 2003, 2005, 2007, 2009)) +
  labs(title = "Annual Counts of Salmon Species",
       x = "Year",
       y = "Amount of Species",
       fill = "Salmon Species",
       caption = "Figure 3. Annual amount of salmon species to use 
       the fish ladder in XX (CITATION HERE).")
```
 



